---
title: "Metagenomics_species"
author: "Sharok"
date: "2023-09-11"
output: html_document
---

```{r}

library(tidyverse)
library(tidytext)

```


```{r}

mapfile <- read.csv("../data/UCFMT1_METAmap.txt", sep = "\t")

```


## Species-level analysis
Reading metaphlan4 default output (merged)
relative abundance
```{r}

org <- read.csv("../data/Metaphlan4/merged_metaphlan4.txt", 
                sep = "\t", header = F, 
                comment.char = "#")
org %>%
  filter(V1 != "clade_name") %>%
  # selecting only at strain level- contains all lineage- avoid duplicates
  filter(grepl("t__S", V1)) -> metaphlan
colnames(metaphlan) <- gsub(".mp.profile_sp", "", org[1,])

metaphlan %>%
  mutate(temp=clade_name) %>%
  separate(temp, c("Kingdom", "Phylum", "Class", 
                       "Order", "Family", "Genus", 
                       "Species", "marker"), sep = "[|]") %>%
  select(clade_name,Kingdom:marker) %>%
  mutate(sp_marker= gsub("t__", "", marker)) %>%
  mutate(sp_marker=gsub("_group", "", sp_marker))-> lineage

```


## engraftment cut-off for species

```{r}

metaphlan %>%
  gather(sample, abundance, -clade_name) %>%
  #filtering donor B markers
  filter(str_detect(sample, "DonorB_D_")) %>%
  #any marker with abundance > 0
  #markers that were present in at least one donor B sample
  filter(!is.na(abundance)) %>%
  filter(as.numeric(abundance) > 0) %>%
  mutate(sample= gsub("DonorB_D_", "", sample)) -> donorB_species

# a function to filter species above cutoff:
Fpos_finder2 <- function(table, cutoff){
  # total number of unqiue species in each sample
  table %>%
    group_by(sample) %>%
    summarise(total=n()) -> total.species
  
  # total number of unique marker above cutoff
  table %>%
    mutate(relative_abund= abundance) %>%
    filter(relative_abund >= cutoff) %>%
    group_by(sample) %>%
    summarise(pos_n=n()) %>%
    left_join(total.species) %>%
    mutate(pos_p=pos_n/total * 100) %>%
    mutate(cutoff=paste(cutoff))
}

cols <- c("2013" = "#bae4b3",
          "2016" = "#74c476", 
          "May17" = "#31a354",
          "Oct17" = "#006d2c")

# comparison of detection cut-offs
Fpos_visualizer2 <- function(table) {
  
  vector <- c(0.001, 0.005, 0.01, 0.1, 1, 2, 3)
  vector <- as.character(vector)  # Convert vector to character for proper factor levels
  
  res <- list()
  for (i in vector) {
    res[[i]] <- Fpos_finder2(table, as.numeric(i))
  }
  
  tbl <- do.call(rbind.data.frame, res)
  tbl$cutoff <- factor(tbl$cutoff, levels = vector)
  
  ggplot(tbl, aes(cutoff, pos_p, color = sample, fill=sample)) +
    geom_line(aes(group = sample)) +
    scale_color_manual(values = cols) +
    scale_fill_manual(values = cols) +
    theme_bw() +
    theme(legend.position = c(0.8,0.7),
          legend.title = element_blank()) +
    xlab("Relative abundance (%)") +
    ylab("% of unique species")
}

p1 <- Fpos_visualizer2(donorB_species)

p2 <- donorB_species %>%
  ungroup() %>%
  group_by(sample) %>%
  summarise(percent=as.numeric(abundance)/sum(as.numeric(abundance)) * 100) %>%
ggplot(aes(x=percent, fill=sample, color=sample)) +
    geom_histogram(position="identity", colour="grey40", alpha=0.7, bins = 10) +
    scale_x_log10() +
    facet_grid(. ~ sample) +
    geom_vline(xintercept = 0.005, color = "red") +
    scale_fill_manual(values = cols) +
    theme_classic() +
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90)) +
  xlab("Relative abundance (%)") +
  ylab("# of species")

cowplot::plot_grid(p1, p2, ncol = 1)
ggsave("../figs/reads_species_cutoff.png",
       width = 14, height = 14, units = "cm")

```

```{r}

metaphlan %>%
  gather(sample, abundance, -clade_name) %>%
  #filtering donor B markers
  filter(str_detect(sample, "DonorB_D_")) %>%
  #any marker with abundance > 0
  #markers that were present in at least one donor B sample
  filter(!is.na(abundance)) %>%
  filter(as.numeric(abundance) > 0) %>%
  mutate(sample= gsub("DonorB_D_", "", sample)) -> donorB_species

mapfile %>% select(Study_ID, Treatment, Timepoint) -> map

##### seting engraftment cut-off
# based on donor B benchmarking data: 0.005 (as shown above)
# minimum relative abundance detected maximum (top quartile) of species
cutoff= 0.005
######

# working with only donor B species that meeet cutoff
# > average of minimum detections
donorB_species %>%
  filter(as.numeric(abundance) > cutoff) -> donorB_species


donorB_species %>%
  select(clade_name) %>% distinct() %>%
  pull() -> B_species

```


## a standard input data for the functions

```{r}

metaphlan %>%
  rename(Marker = "clade_name") -> marker_lvl_RA


#write.csv(marker_lvl_RA, "../data/species_lvl_RA.csv",
#          quote = F, row.names = F)

```

## Functions

```{r}
# Make the marker data frame long
mark_to_long = function(marker_lvl, strains) {
    # This can be done once. It doesn't need to be repeated each time
    # convert marker-lvl to long format and make StudyID variable to match mapfile
    # and select only donor's strains
    marker_lvl %>%
        filter(Marker %in% strains) %>%
        gather(sample, abundance, -Marker) %>%
        mutate(Study_ID = case_when(
            str_detect(sample,"PMCL") ~ paste(gsub("_.*", "",sample)),
            TRUE ~ paste(sample))) -> long_markerlvl
    return(long_markerlvl)
}

################################################################################
# a function to detect common engraftment based on strains (metaphlan markers)
################################################################################
get_engraft <- function(long_markerlvl, mapfile, cutoff=cutoff, ...){
  # complete list of engrafted
  engraft <- (long_markerlvl 
              %>% left_join(mapfile) 
              %>% filter(...)
              %>% spread(Timepoint, abundance, fill = 0) 
              %>% filter(WK0 <= 0 & WK6 > cutoff) 
              %>% select(Marker, Study_ID, WK6) 
              %>% spread(Study_ID, WK6, fill = 0) 
              %>% gather(Study_ID, WK6, -Marker)
              %>% left_join(mapfile))
    return(engraft)
}

count_engraft <- function(engraft, group_cols){
  # Count how many markers engrafted in how many patients
  g1 = group_cols
  g2 = c('Marker', group_cols)
  ct_df = (engraft
           %>% group_by(.[,g2])
           %>% summarize(n = sum(WK6 > 0))
           %>% pivot_wider(names_from = {{ g1 }},
                           values_from = n)
           %>% ungroup()
           %>% mutate(Total = rowSums(across(2:ncol(.)))))
    # Uncomment the next bit if you want the cumulative sums, but it adds 0.8s
    # to the run time!
    # ct_df = (ct_df
    #          %>% mutate(n = rev(cumsum(rev(n))))) 
    return(ct_df)
    
}

################################################################################


```


### Observed AUC: FMT vs Placebo

```{r}

# Remove unnecessary columns from the mapfile to speed up the loop (saves 10s per 
# loop)
map_mod = (mapfile
           %>% select(Study_ID, Timepoint, Treatment, Fig_lab))

# Converting the markers to long format only needs to be done once, saves 5s per
# loop
markers_long = mark_to_long(marker_lvl_RA, B_species)

cts = (markers_long
       %>% get_engraft(map_mod, cutoff, 
                       Treatment %in% c('FMT', 'Placebo'))
       %>% count_engraft(group_cols = 'Treatment')
       %>% mutate(Uniqueness = 
                    case_when(FMT == 0 & Placebo == 0 ~ 'zero',
                              FMT == 0 ~ 'Placebo',
                              Placebo == 0 ~ 'FMT',
                              TRUE ~ 'Both'),
                  it = 'observed'))
npts_tx = rbind(
  (cts
   %>% count(FMT)
   %>% rename(Npts = FMT)
   %>% mutate(Treatment = 'FMT')),
  (cts
   %>% count(Placebo)
   %>% rename(Npts = Placebo)
   %>% mutate(Treatment = 'Placebo')))

fmt_fn = with(filter(npts_tx, Treatment == 'FMT'),
              approxfun(Npts, n))
plc_fn = with(filter(npts_tx, Treatment == 'Placebo'),
               approxfun(Npts, n))
fmt_auc = with(filter(npts_tx, Treatment == 'FMT'),
                   integrate(fmt_fn, 1, max(Npts))$value)
plc_auc = with(filter(npts_tx, Treatment == 'Placebo'),
                   integrate(plc_fn, 1, max(Npts))$value)
obs_d_auc = fmt_auc - plc_auc

obs_d_auc
```


### Get observed AUC p-value: FMT vs Placebo
loading permuated data "auc_null"

```{r}

# load data:
auc_null<- read.csv("../permut_data/species_auc_null.txt",
                    sep = "\t")

# Get the observed FMT values
ggplot(data.frame(auc_null), aes(x = deltaAUC)) +
    geom_density() +
    geom_vline(xintercept = obs_d_auc) +
  theme_classic()

# Place that difference on the null distribution of differences and get its p-value
quant_fun = ecdf(auc_null[,1])
auc_pval = 1 - quant_fun(obs_d_auc)
auc_pval

# merging the results together:
plt_df = filter(npts_tx, Npts > 0)
ribdat = plt_df %>% spread(Treatment, n, fill = 0)
  
cols <- c("FMT" = "#1b7837",
          "Placebo" = "#999999")

f1 <- ggplot(plt_df, aes(Npts, n, color=Treatment)) +
  geom_point() + geom_line() +
  geom_ribbon(data = ribdat, aes(ymin=Placebo,ymax=FMT, x = Npts),
              linetype=2, alpha=0.1, inherit.aes = FALSE)+
  annotate("text", x=9, y=90, size=2,
          label= paste("Permutation test,\n p=", auc_pval, sep = " ")) +
  scale_x_continuous(breaks = seq(0, max(plt_df$Npts), 1)) +
  scale_color_manual(values = cols) +
  theme_classic() +
  theme(legend.position = "none",
        legend.title = element_blank(),
        axis.title = element_blank())
```

## Uniqueness
test whether the engrafted markers within FMT and Placebo are shared between
these groups or unique to each treatment

```{r}

plt2_df = (cts
           %>% count(Total, Uniqueness)
           %>% group_by(Total))


cols <- c("FMT" = "#1b7837",
          "Placebo" = "#999999",
          "Both" = "#fc8d59")

f2 <- ggplot(plt2_df, aes(x = Total, y = n, fill = Uniqueness)) +
  geom_bar(stat = 'identity') +
  scale_fill_manual(values = cols) +
  theme_classic() +
  theme(legend.position = "none",
        axis.title = element_blank()) 

```

### Observed AUC: Res vs No-Res

```{r}

# Remove unnecessary columns from the mapfile to speed up the loop (saves 10s per 
# loop)
map_mod = (mapfile
           %>% filter(Treatment == 'FMT')
           %>% select(Study_ID, Timepoint, Remission, Fig_lab))

fmt_cts = (markers_long
       %>% get_engraft(map_mod, cutoff, 
                       Remission %in% c('Res', 'NoRes'))
       %>% count_engraft(group_cols = 'Remission')
       %>% mutate(Uniqueness = 
                    case_when(Res == 0 & NoRes == 0 ~ 'zero',
                              Res == 0 ~ 'NoRes',
                              NoRes == 0 ~ 'Res',
                              TRUE ~ 'Both'),
                  it = 'observed'))
npts_fmt = rbind(
  (fmt_cts
   %>% count(Res)
   %>% rename(Npts = Res)
   %>% mutate(Remission = 'Res')),
  (fmt_cts
   %>% count(NoRes)
   %>% rename(Npts = NoRes)
   %>% mutate(Remission = 'NoRes')))

res_fn = with(filter(npts_fmt, Remission == 'Res'),
              approxfun(Npts, n))
nrs_fn = with(filter(npts_fmt, Remission == 'NoRes'),
               approxfun(Npts, n))
res_auc = with(filter(npts_fmt, Remission == 'Res'),
               sum(integrate(res_fn, 0, 1)$value,
                   integrate(res_fn, 1, max(Npts))$value))
nrs_auc = with(filter(npts_fmt, Remission == 'NoRes'),
               sum(integrate(nrs_fn, 0, 1)$value,
                   integrate(nrs_fn, 1, max(Npts))$value))
rm_d_auc = res_auc - nrs_auc


```

### Get observed AUC p-value: Res vs No-Res
loading permutated data "rm_auc_null"

```{r}

# load data:
rm_auc_null <- read.csv("../permut_data/species_remission_auc_null.txt",
                    sep = "\t")

# Get the observed FMT values
ggplot(data.frame(rm_auc_null), aes(x = deltaAUC)) +
    geom_density() +
    geom_vline(xintercept = rm_d_auc) +
  theme_classic()

# Place that difference on the null distribution of differences and get its p-value
quant_fun = ecdf(rm_auc_null[,1])
auc_pval = 1 - quant_fun(rm_d_auc)
auc_pval

# merging the results together:
plt_df = filter(npts_fmt, Npts > 0)
ribdat = plt_df %>% spread(Remission, n, fill = 0)


cols <- c("Res" = "#2166ac",
          "NoRes" = "#b2182b")
  
f3 <- ggplot(plt_df, aes(Npts, n, color=Remission)) +
  geom_point() + geom_line() +
  geom_ribbon(data = ribdat, aes(ymin=NoRes,ymax=Res, x = Npts),
              linetype=2, alpha=0.1, inherit.aes = FALSE)+
  scale_x_continuous(breaks = seq(0, max(plt_df$Npts), 1)) +
  annotate("text", x=5, y=90, size=2,
           label= paste("Permutation test,\n p=", auc_pval, sep = " ")) +
  scale_color_manual(values = cols) +
  theme_classic() +
  theme(legend.position = "none",
        axis.title = element_blank())

```

### saving figure
```{r}

cowplot::plot_grid(f1, f2, f3, ncol = 1, hjust = T, vjust = T)
ggsave("../figs/reads_CESp.png",
       width = 6, height = 12, units = "cm")

```